---
title: "Global population concentration"
author: "Piyush Verma"
date: "December 27, 2017"
output: html_document
---


* Coastal China and Indian subcontinent region are the most densely populated areas in the world
* Though Tokyo in Japan has the biggest population
* North American population is majorly centred around the North-eastern United States
* Aussies, as going by their popularity of being beach bums, live mostly on the coast (their innermost land is basically uninhabitable)
* Regions like Sahara, Gobi desert, Greenland, Amazon forest and Siberia are scarcely inhabited.


```{r setup, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8)
```


```{r, message=FALSE,warning=FALSE,results='hide', echo=FALSE}
set.seed(2017-12-27)
library("data.table")
cities<-fread("./worldcities.csv")
cities<-cities[cities$pop>0,]
```

```{r, message=FALSE,warning=FALSE, echo=FALSE, width = 40, height = 30}
library("leaflet")
pal <- colorNumeric(palette = "Red",domain = cities$pop)
cities2<-cbind(cities,col=pal(cities$pop))

my_map <- cities2 %>% leaflet() %>% addTiles() %>% addCircles(weight = 1, radius = sqrt(cities$pop) * 110) %>% setView(lat = 51.4826, lng = 0.0077, zoom = 2) 
my_map
```



